EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-08-13 draft.tex
Rewrite of section 6
I describe the aws setup used in the following section. @jj  please have a look.

EvoSwarm describes the deployment of loosely coupled processes in a consistent
and portable way. The term cloud-native describes these container-based
environments because we can deploy them on an infrastructure that abstracts the
underlying compute, storage, and networking primitives. This infrastructure does
not need to be a cloud provider; it could be a developer's laptop, a single
server, or a production environment. In this section, we answer the following
question:

Can EvoSwarm as a cloud-native, container-based application be deployed locally
or in the cloud by specifying the configuration of container images and
resources in an orchestration service?

To prove this question, we present a brief optimization experiment that is
deployed in three distinct environments, taking advantage of it as a
container-based application. When we deploy a cloud-native application to a cloud provider, we have some more
options to choose:

@@

  \item[A virtual machine] with a container-optimized operating system, for example,
  Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
  virtual machine and run {\tt docker-compose} commands as we would do in a local 
  workstation. Salza et al. \cite{salza2019speed} tested this option by using
  CoreOs cluster.

@@

  \item[A container orchestration service] these services are compatible with {\tt docker-compose}
    files for configuration and provide transparent provisioning. Examples of these
    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.

@@

We chose the third option, that is, a container orchestration service, and
deployed the EvoSwarm application in Amazon Elastic Container Service (ECS),
since it needs fewer changes in configuration (from deploying it locally using
{\tt docker-compose}) and it is has a lower cost than the Kubernetes
alternative.

% Remember to mention cost at the end - JJ
% Ping - JJ
We can define an ECS cluster directly on AWS Elastic Cloud VMs or through the AWS Fargate service.
Fargate is a serverless compute engine for containers that provides automatic provision 
and management of AWS ECS resources. With this service, calculating the cost of execution
is simplified because users only specify the computing units and memory for all the application,
instead of every component.

Multicontainer applications are deployed to ECS as tasks, and a single cluster can run multiple tasks in
parallel. When using the Fargate service as a backend, tasks are limited to a
maximum of 10 containers.

Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can specify the configuration of a task 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt task\_size} indicating the memory and CPU units
assigned, and network configuration. The complete configuration is generated
by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
the configuration on-the-fly. For example to deploy the controller, setup and Redis 
containers we used the file in Listing ~\ref{code:compose}.
Each container is defined as a  {\texttt service} , with the following attributes: {\texttt image},
that reference the container images we have published in Docker hub
\url{https://hub.docker.com/u/mariosky}, open {\texttt ports}, a {\texttt command} to be executed at
startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}
variables used by code running inside containers.

We deployed two ECS configurations, one that included all containers in a single task,
and another one, where we assigned each worker to a different task and the controller, 
Redis, and setup in another, for a total of nine concurrent Fargate tasks.

For the first option, we deployed an AWS Fargate cluster and defined a task configuration file with
8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
ten container limit, we only used seven workers and tested function $f_4$ as
in the previous experiment, and to save resources we only tested for dimensions 10 and 20 for 15 instances.

For the second option, we used the same cluster and additionally
assigned each worker to a task with 4GBs of memory and 2048 computing units. Then scale the 
number of worker tasks to eight. We can scale the number of tasks for up to 100 instances in 
a Fargate configuration.

For local experiments, we have deployed the EvoSwarm application in a
Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and {\tt docker-compose} version 1.21.0, 
in Ubuntu Linux 18.04; code is run using Python 3.7.5. Container images and the {\tt docker-compose} file are available at
\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively.

We reproduced the experiment of Listing~\ref{code:exp} in each environment. The experiment consists of
solving $f_4$ for 10 and 20 dimensions. We give more details about additional parameters of the configuration in Section~\ref{setup}.

\begin{lstlisting}[caption=ecs-params.yml , label=code:ecs]
version: 1
task_definition:
  task_execution_role: ecsTaskExecutionRole
  ecs_network_mode: awsvpc
  task_size:
    mem_limit: 8GB
    cpu_limit: 4096
\end{lstlisting}

@@

In this experiment, we wanted to test several deployment options for the
cloud-native application, and have an idea of the computing capacity of each
option measured as the number of function evaluations per second. The results
are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}. As we can
see, the local environment (see section \ref{setup}) achieved a higher ratio of
evaluations per second (in bold). Still, we must consider that there is minimum
network overhead in a local workstation and a clock speed of 3.8 GHz. The first
option (labeled Single) is straightforward to deploy, needing only a single {\tt
docker-compose} file, and one command to execute. But with the limit of 4 CPUs
and ten containers, it is the slowest option. The second option (labeled Multi),
is faster, and not too far from the local execution, especially when solving the
$f_4$ in 20 dimensions (underlined in table). With the capacity for increasing
the number of worker containers, it is reasonable to think that we can achieve a
higher number of evaluations with an appropriate configuration. Nevertheless,
the goal of this experiment was not to increase the number of evaluations but to
test the flexibility of a cloud-native application. The total cost of testing,
configuration, and execution of the experiments in ECS was less than \$1.00 US
Dollar.

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{aws}
  \caption{Boxplot of the number of evaluations per second, for 15 instances of function $f_4$,
   with three execution environments Local, Single task, and Multi-task.
   }
  \label{fig:aws-results}
\end{figure}
%
\begin{table}[h!tbp]
  \small
  \caption{ Median number of the evaluations per second, for 15 instances of function $f_4$,
  with three execution environments Local, Single task, and Multi task. The best result is shown in {\bf boldface}.}
  \label{tab:aws}
  \vspace{0.25cm}
  \centering

  \begin{tabular}{lll}
  \hline
  Dim & Env         & Median of ratio \\
  \hline
  10  & Local       & \textbf{47,166.29}       \\
      & Multi task  & 40,466.45    \\
      & Single task & 18,193.05      \\
  \hline
  20  & Local       & \textbf{36,101.86}   \\
      & Multi task  & {\ul 33,916.02}       \\
      & Single task & 16,546.95  \\
  \hline
  \end{tabular}
\end{table}

The same workflow and configuration files used by this brief experiment, will be
used on the next section, to carry a more extensive evaluation of the framework.
Again, each configuration describe the deployment of containers in a consistent
and portable way, an it was used in each phase of the life cycle, from
development through testing to production independently on the kind of
infrastructure.

@@

In the prior section, we wanted to prove that EvoSwarm can be used by 
the public to create reproductive research, but in this section are
using the same framework to learn if the proposed solution can efficiently
improve the scalability and performance of population-based optimization
algorithms. % If you want to prove it *improves*, you need to provide a baseline. 
            %What will be your baseline? You need to specify anyway 
            %in which way it really improves - JJ
            % The baseline is a single population GA or PSO? - M
% For instance. But I was actually looking for another word that does not imply a comparison - JJ
Hence, in the following sections, we verify the following questions:

>>>>

EvoSwarm describes the deployment of loosely coupled processes in a consistent
and portable way. The term cloud-native describes these container-based
environments because we can deploy them on an infrastructure that abstracts the
underlying compute, storage, and networking primitives. This infrastructure does
not need to be a cloud provider; it could be a developer's laptop, a single
server, or a production environment. In this section, we give specific details 
about how we deployed  EvoSwarm in a cloud-setting for running the experiments 
within this paper.

When we deploy a cloud-native application to a cloud provider, we have two
options to choose:

@@

@@

  \item[A container orchestration service] these services are compatible with {\tt docker-compose}
    files for configuration and provide transparent provisioning. Examples of these
    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.
    Salza et al. \cite{salza2019speed} tested this option by using CoreOs cluster.

@@

We chose the second option, that is, a container orchestration service, and
deployed the EvoSwarm application in Amazon Elastic Container Service (ECS),
since it needs fewer changes in configuration (from deploying it locally using
{\tt docker-compose}) and it is has a lower cost than the Kubernetes
alternative.

% Remember to mention cost at the end - JJ 
% Ping - JJ

We can define an ECS cluster directly on the AWS Elastic Cloud VMs or through the
AWS Fargate service. Fargate is a serverless compute engine for containers that
provides automatic provision and management of AWS ECS resources. With this
service, calculating the cost of execution is simplified because users only
specify the computing units and memory for the application, instead of every
component. However, AWS Fargate is not suited for an experimental setting because
many of the details are out of user control. For this reason, we chose the
alternative of using a cluster of EC2 containers. The drawback is that we need to manually
provision the  EC2 containers, networking, and group permissions for each
experimental setup, for instance, when going from 2 to 8 workers, we need to increase the
number of EC2 instances to meet the required amount of CPU units.

Multicontainer applications are deployed to ECS as tasks, and a single cluster can run multiple tasks in
parallel. Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can also specify the configuration of a task 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt cpu\_shares} indicating the CPU units
assigned, in this case 2048; and the network configuration, we have two private 
subnets and a security group in this configuration. When deploying to EC2 containers, these requirements
need to be supported by the currently available cluster. For instance, we need to have all 
the CPU units required by all tasks, if they are not available, the task initiation will fail.

The complete task configuration is generated by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
the configuration on-the-fly. For example to deploy the controller, setup and Redis 
containers we used the file in Listing ~\ref{code:compose}.

Each container is defined as a {\texttt service} (task in AWS terms) , with the following attributes: {\texttt image},
that reference the container images we have published in Docker hub
\url{https://hub.docker.com/u/mariosky}, open {\texttt ports}, a {\texttt command} to be executed at
startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}
variables used by code running inside containers.

Figure \ref{fig:aws-configuration} shows the basic EC2 configuration used for the deployment
of experiments. We selected three types of EC2 instances according to our needs, the  
{\texttt setup}, {\texttt controller}, and {\texttt redis} containers do not demand many resources
and we do not need to scale them as the amount of parallel workers are increased. They run in a
single {\texttt C5  High-CPU large} instance, with the available 2048 CPU shares divided as follows:
{\texttt setup} (128),  {\texttt controller} (896),{\texttt redis} (1024). We assigned 2048 units to 
each {\texttt worker}   container, by using a more capable {\texttt C5 High-CPU XLarge} instance,  with 
four vCPUs and 4096 cpu shares. We scaled the number if instances as needed, using one instance
for every two {\texttt worker} containers. Because all these instances run in a private subnet, we
run a Bastion instance in a public subnet, from where we executed the experiments and collected the results,
through a {\texttt ssh} session. This is a {\texttt R5AD Large} instance,  with 75GiB of storage for
storing the logs, and more memory (16GiB) for processing the logs. The details and cost of these instances at the 
time of writing are shown in Table \ref{tab:ec2}.

\begin{table}[h!tbp]
  \small
  \caption{Details of EC2 instance types used in our deployment}
  \label{tab:ec2} 
  \centering
  \small
  \begin{tabular}{|l|l|r|l|l|}
    \hline
    Name & Memory & vCPUs & Storage  & Cost \\ \hline
    R5AD Large	&	16.0 GiB	& 2 vCPUs	& 75 GiB 	& \$0.131000 hourly \\ \hline
    C5 High-CPU Large	&	4.0 GiB	& 2 vCPUs	& EBS Only 	& \$0.102000 hourly \\ \hline
    C5 High-CPU XLarge	&	8.0 GiB	& 4 vCPUs	& EBS Only 	& \$0.154000 hourly \\ \hline
  \end{tabular}
\end{table}

\begin{lstlisting}[caption=ecs-params.yml , label=code:ecs]
  version: 1
  task_definition:
    ecs_network_mode: awsvpc
    services:
      worker:
        cpu_shares: 2048

  run_params:
    network_configuration:
      awsvpc_configuration:
        subnets:
          - "subnet-0ca2a542bbd74239e"
          - "subnet-0dfd367b5ca531b07"
        security_groups:
          - "sg-0681c18469874b391"  
\end{lstlisting}

@@

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{aws_ec2}
  \caption{AWS EC2 configuration for the deployment of EvoSwarm containers.
   }
  \label{fig:aws-configuration}
\end{figure}
%

This configuration files will be
used on the next section, to carry an extensive evaluation of the framework.
Again, each configuration describes the deployment of containers in a consistent
and portable way, an it was used in each phase of the life cycle, from
development through testing to production independently on the kind of
infrastructure.

@@

In the prior section, we describe an option for an EvoSwarm deployment that can be used by 
the public to create reproductive research, in this section we are
using the same framework to learn if the proposed solution can efficiently
improve the scalability and performance of population-based optimization
algorithms. % If you want to prove it *improves*, you need to provide a baseline. 
            %What will be your baseline? You need to specify anyway 
            %in which way it really improves - JJ
            % The baseline is a single population GA or PSO? - M
% For instance. But I was actually looking for another word that does not imply a comparison - JJ
Hence, in the following sections, we verify the following questions:
